When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph Databases

### 2018-02-23 paper/paper.tex
Calls per second
Bit of a reminder for non-CP people

Our experiments will primarily measure the number of \emph{search nodes} (that is, the number of
recursive calls made, or the number of guessed assignments), not runtimes. We are not aiming to give
a simple comparison of absolute performance between solvers; rather, we are looking for
solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which is
enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers experienced many more
failures with this timeout, so our picture of just how hard the hardest instances are with these
solvers is less detailed.

@@

Various general principles have been considered when designing variable- and value-ordering
heuristics for backtracking search algorithms---one of these is to try to maximise the expected
number of solutions inside any subproblem considered during search \citep{DBLP:conf/cp/GentMPSW96}.
This is usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD
and Glasgow pick a variable with fewest remaining values in its domain: doing this will generally
reduce the first part of the $\langle Sol \rangle$ equation (i.e.\ $t^{\underline{p}}$) by as little
as possible.  When two or more domains are of equal size, LAD breaks ties lexicographically, whereas
Glasgow will pick a variable corresponding to a pattern vertex of highest degree. This strategy was
determined empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking
a pattern vertex of high degree will make the remaining pattern subgraph sparser, which will
decrease the exponent in the second half of the formula, maximising the overall value. LAD does not
apply a value-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.
Again, this was determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by
leaving as many vertices as possible available for future use. The search strategy used by VF2, in
contrast, is based around preserving connectivity, which gives very little discrimination except on
the sparsest of inputs.

>>>>

Our experiments will primarily measure the number of \emph{search nodes} (that is, the number of
recursive calls made, or the number of guessed assignments), not runtimes. We are not aiming to give
a simple comparison of absolute performance between solvers; rather, we are looking for
solver-independent patterns of difficulty. However, roughly speaking, on the small graphs used in
the following sections, Glasgow manages around 300,000 recursive calls per second on our hardware,
whilst LAD and VF2 perform around 1,000 and 70,000 recursive calls per second respectively.
All experiments use a timeout of 1,000 seconds, which is enough for the Glasgow solver to solve
nearly all instances (whose orders were selected with this timeout in mind), although we may
slightly overestimate the proportion of unsatisfiable instances for extremely sparse or dense
pattern graphs. The LAD, VF2, and VF3 solvers experienced many more failures with this timeout, so
our picture of just how hard the hardest instances are with these solvers is less detailed.

@@

The Glasgow and LAD solvers are driven by ordering heuristics: variable-ordering tells us which
pattern vertex to branch on, whilst value-ordering tells the order in which we should try target
vertices for that pattern vertex.
Various general principles have been considered when designing variable- and value-ordering
heuristics for backtracking search algorithms---one of these is to try to maximise the expected
number of solutions inside any subproblem considered during search \citep{DBLP:conf/cp/GentMPSW96}.
This is usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD
and Glasgow pick a variable with fewest remaining values in its domain: doing this will generally
reduce the first part of the $\langle Sol \rangle$ equation (i.e.\ $t^{\underline{p}}$) by as little
as possible.  When two or more domains are of equal size, LAD breaks ties lexicographically, whereas
Glasgow will pick a variable corresponding to a pattern vertex of highest degree. This strategy was
determined empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking
a pattern vertex of high degree will make the remaining pattern subgraph sparser, which will
decrease the exponent in the second half of the formula, maximising the overall value. LAD does not
apply a value-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.
Again, this was determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by
leaving as many vertices as possible available for future use. The search strategy used by VF2, in
contrast, is based around preserving connectivity, which gives very little discrimination except on
the sparsest of inputs.
